MS_Intro_Short

### 2022-05-06 Markdown/Manuscript/Child_RMDs/Doodles.Rmd
minor tweaks

>>>>

In the limited cases where code is made publicly available it is typically in the form of programs (an algorithm that performs one task) rather than code libraries (where the primary algorithm is decomponsed into multiple component algorithms, that can be re-used for a potentially large number of other purposes).

### 2022-05-06 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
minor tweaks

<!-- `r ifelse(is.na(params$X@background_1L_chr),"",params$X@background_1L_chr)`  -->
Mental disorders impose high health, social and economic burdens worldwide [@RN8][@GBD2019]. Much of this burden is potentially avertable [@RN25], but poorly financed and organised mental health systems are ill-equiped for this challenge[@RN22][@RN23]. The large and widespread additional mental health burdens recently observed during the COVID pandemic[@20211700] and predicted as a potential future consequence of global heating [@page_howard_2010], highlight the need to improve the resilience and adaptability of these systems. To help stem growing demand for mental health services, policymakers have also been encouraged to place greater emphasis on tackling the social determinants of mental disorder[@RN11].

@@

Open source frameworks, which have been recommended for the development of mental health modelling field (Long, 2018), have the potential to help address a number of these challenges. Open source approaches can promote common standards and interoperability, improve model transparency and facilitate routine updates to rectify errors, update data and add features. They may also have educational value in developing the health systems modelling field (Dunlop, 2017).

However, open source computational models also present a range of conceptual, technical, legal and resourcing issues compared to more traditional approaches to model development.

However, of the three components that comprise a mental health computer simulation model - the conceptual model, the mathematical model and the computational model - only the first two are typically developed and reported in a manner that facilitates reuse, refinement and extension by others. This is problematic as the code implementation is where the theory represented by a simulation model is most fully elaborated. In the limited cases where code is made publicly available it is typically in the form of programs (an algorithm that performs one task) rather than code libraries (where the primary algorithm is decomponsed into multiple component algorithms, that can be re-used for a potentially large number of other purposes).

>>>>

<!-- `r ifelse(is.na(params$X@background_1L_chr),"",params$X@background_1L_chr)`  -->
Mental disorders impose high health, social and economic burdens worldwide [@RN8;@GBD2019]. Much of this burden is potentially avertable [@RN25], but poorly financed and organised mental health systems are ill-equiped for this challenge[@RN22;@RN23]. The large and widespread additional mental health burdens recently observed during the COVID pandemic[@20211700] and predicted as a potential future consequence of global heating [@page_howard_2010], highlight the need to improve the resilience and adaptability of these systems. To help stem growing demand for mental health services, policymakers have also been encouraged to place greater emphasis on tackling the social determinants of mental disorder[@RN11].

@@

Open source frameworks, which have been recommended for the development of mental health modelling field [@RN73], have the potential to help address a number of these challenges. Open source approaches can promote common standards and interoperability, improve model transparency and facilitate routine updates to rectify errors, update data and add features. They may also have educational value in developing the health systems modelling field [@Dunlop2017]. However, open source computational models also present a range of conceptual, technical, legal and resourcing issues compared to more traditional approaches to model development.

- Our project - readyforwhatsnext : open source model of the systems shaping the mental health of young people
- Lack of guidance on open source health economic models and in particular on recommended frameworks for implementing the computational model.
- Focus computational model. Most in need of novel guidance. Also: Of the three components that comprise a mental health computer simulation model - the conceptual model, the mathematical model and the computational model - only the first two are typically developed and reported in a manner that facilitates reuse, refinement and extension by others. This is problematic as the code implementation is where the theory represented by a simulation model is most fully elaborated.
